Type-Safe Modular Parsing

### 2017-01-14 paper/sections/Introduction.tex
Intro

To address the modularity limitations of Programming Languages, several
different approaches have been proposed in the past. Existing
approaches can be broadly divided into two categories:
\emph{syntactic} or \emph{semantic} modularization
techniques. Syntactic modularization techniques are quite popular in
practice, due to their simplicity of implementation and use.
Examples include many tools for developing Feature-Oriented Software-Product
Lines (SPLs)~\cite{Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.  Most syntactic approaches employ textual
composition techniques such as \emph{superimposition}~\cite{Wagelaar2010} to
enable the development modular program features. Such
composition techniques collect the code for multiple features and
merge them together when a concrete combination of features is needed
for a particular program. As Kastner et. al~\cite{Kastner11road} note,
a typical drawback of feature-oriented SPL implementations, which
more generally applies to syntactic modularity approaches, is that
such ``\emph{implementation mechanisms lack proper
  interfaces and support neither modular type checking nor separate
  compilation}''.

@@

Semantic modularization techniques go one step further in terms of modularity,
and also enable components or features to be modularly type-checked
and separately compiled. Modular type-checking and separate
compilation are desirable properties to have from a software
engineering point-of-view. Modular type-checking can report errors
earlier and in terms of the modular code programmers have written
in the first place. Separate compilation avoids global compilation
steps, which can be very costly. Furthermore semantic modularization
enables the composition of compiled binaries as well as ensuring the
type-safety of the code composed of multiple components. Examples of semantic modularization techniques
include various approaches to \emph{family polymorphism}~\cite{ernst01FP},
\emph{virtual classes}~\cite{Ernst:2006}, as
well as various techniques for solving the Expression
Problem~\cite{Oliveira:2012,wang2016expression,torgersen2004expression,odersky2005independently}.
Semantic modularization techniques are less widely used in practice
than syntactic techniques. This is partly due to the perceived need for more
sophisticated type systems, which are not available in mainstream
languages and may require more knowledge from users. However, recently,
several lightweight modularization techniques have been shown to work
in mainstream programming languages like Java or Scala. Object
Algebras~\cite{Oliveira:2012} are one such technique, which works in
Java-like languages and uses simple generics only.

>>>>

To address the modularity limitations of Programming Languages, several
different approaches have been proposed in the past. Existing
approaches can be broadly divided into two categories:
\emph{syntactic} or \emph{semantic} modularization
techniques. Syntactic modularization techniques are quite popular in
practice, due to their simplicity of implementation and use.
Examples include many tools for developing Feature-Oriented Software-Product
Lines (SPLs)~\cite{AK:JOT09,Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.  Most syntactic approaches employ textual
composition techniques such as \emph{superimposition}~\cite{AK:JOT09} to
enable the development modular program features. Such
composition techniques collect the code for multiple features and
merge them together when a concrete combination of features is needed
for a particular program. As Kastner et. al~\cite{Kastner11road} note,
a typical drawback of feature-oriented SPL implementations, which
more generally applies to syntactic modularity approaches, is that
such ``\emph{implementation mechanisms lack proper
  interfaces and support neither modular type checking nor separate
  compilation}''.

@@

Semantic modularization techniques go one step further in terms of modularity,
and also enable components or features to be modularly type-checked
and separately compiled. Modular type-checking and separate
compilation are desirable properties to have from a software
engineering point-of-view. Modular type-checking can report errors
earlier and in terms of the modular code programmers have written
in the first place. Separate compilation avoids global compilation
steps, which can be very costly. Furthermore semantic modularization
enables the composition of compiled binaries as well as ensuring the
type-safety of the code composed of multiple components. Examples of semantic modularization techniques
include various approaches to \emph{family polymorphism}~\cite{ernst01FP},
\emph{virtual classes}~\cite{Ernst:2006}, as
well as various techniques for solving the Expression
Problem~\cite{torgersen2004expression,odersky2005independently,Oliveira:2012,wang2016expression}.
Semantic modularization techniques are less widely used in practice
than syntactic techniques. This is partly due to the perceived need for more
sophisticated type systems, which are not available in mainstream
languages and may require more knowledge from users. However, recently,
several lightweight modularization techniques have been shown to work
in mainstream programming languages like Java or Scala. Object
Algebras~\cite{Oliveira:2012} are one such technique, which works in
Java-like languages and uses simple generics only.
